\section{2-Server Stateless PIR via DPF}
\label{sec:2server}

This section describes the 2-server stateless PIR instantiation of our system, based on Distributed Point Functions (DPF). The client requires no persistent state or hints---only the DPF keys generated fresh for each query session. Combined with the batch PIR framework from Section~\ref{sec:batching}, this yields a complete system for private UTXO lookups.

\subsection{Distributed Point Functions (DPF)}
\label{sec:dpf}

A Distributed Point Function (DPF) \cite{gilboa2014dpf,boyle2015fss,boyle2016fss} is a cryptographic primitive that enables efficient 2-server PIR. A \emph{point function} $f_{\alpha,\beta}$ evaluates to $\beta$ at input $\alpha$ and to $0$ everywhere else. A DPF splits this point function into two keys $(k_0, k_1)$ such that:
\begin{itemize}[nolistsep,leftmargin=*]
    \item \textbf{Correctness:} $\text{Eval}(k_0, x) \oplus \text{Eval}(k_1, x) = f_{\alpha,\beta}(x)$ for all $x$.
    \item \textbf{Security:} Each individual key $k_b$ is computationally indistinguishable from a key for any other point function.
\end{itemize}

Each key has size $O(\lambda \log N)$ where $\lambda$ is the security parameter and $N$ is the domain size. To perform PIR on a database of $N$ entries indexed by $[N]$:
\begin{enumerate}[nolistsep,leftmargin=*]
    \item The client generates DPF keys $(k_0, k_1)$ for the point function $f_{i, 1}$ where $i$ is the target index.
    \item Server $b$ expands $k_b$ to obtain a vector of $N$ field elements and computes $r_b = \bigoplus_{x=0}^{N-1} \text{Eval}(k_b, x) \cdot D[x]$.
    \item The client recovers $D[i] = r_0 \oplus r_1$.
\end{enumerate}

The server computation is $O(NB)$ but consists of simple XOR operations over the database, making it extremely fast in practice---often close to the speed of a single linear pass over the data in memory.

\subsection{Implementation}
\label{sec:implementation}

We have implemented a complete system based on the 2-server DPF-PIR approach with Probabilistic Batch Codes (PBC), as described in Section~\ref{sec:batching}. The implementation consists of database generation tools (Rust), two PIR servers (Rust with \texttt{tokio}), and client libraries (Rust CLI and TypeScript/Web). The two-level architecture from Figure~\ref{fig:two-maps} eliminates the TXID resolution step of earlier designs by storing full 32-byte TXIDs directly in the UTXO chunks.

\subsubsection{System Architecture}

Each server maintains identical copies of the index and chunk databases, loaded via memory-mapped files (\texttt{memmap2}). The client connects to both servers over WebSocket and distributes DPF key shares, so neither server learns the query target unless they collude. Communication uses a compact binary protocol supporting database discovery, batch PIR queries, and connection health checks.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    node distance=1.5cm and 2cm,
    box/.style={rectangle, draw, rounded corners, minimum width=3cm, minimum height=1.2cm, align=center, font=\small},
    db/.style={cylinder, draw, shape border rotate=90, aspect=0.25, minimum height=1.5cm, minimum width=1.2cm, fill=blue!10, font=\footnotesize},
    arrow/.style={-{Stealth[length=2.5mm]}, thick},
    label/.style={font=\scriptsize, text=gray!70!black}
]

\node[box, fill=green!20] (client) {Client\\{\scriptsize (Rust CLI / Web)}};
\node[box, fill=orange!20, below left=1.5cm and 1cm of client] (server1) {Server 1};
\node[box, fill=orange!20, below right=1.5cm and 1cm of client] (server2) {Server 2};
\node[db, below=0.5cm of server1] (db1) {Index + Chunks};
\node[db, below=0.5cm of server2] (db2) {Index + Chunks};

\draw[arrow] (client) -- (server1) node[midway, above, sloped, label] {$K$ DPF keys $\{k_1^{(i)}\}$};
\draw[arrow] (client) -- (server2) node[midway, above, sloped, label] {$K$ DPF keys $\{k_2^{(i)}\}$};
\draw[arrow, dashed, gray] (server1) -- (db1);
\draw[arrow, dashed, gray] (server2) -- (db2);

\draw[arrow, blue!60!black] (server1.north east) -- ++(0.3, 0.5) node[right, font=\scriptsize] {$\{r_1^{(i)}\}$};
\draw[arrow, blue!60!black] (server2.north west) -- ++(-0.3, 0.5) node[left, font=\scriptsize] {$\{r_2^{(i)}\}$};

\node[above=1.8cm of client, font=\small, text=blue!60!black] (xor) {Results: $r_1^{(i)} \oplus r_2^{(i)}$ for $i \in [K]$};
\draw[arrow, blue!60!black, dashed] ($(server1.north east)+(0.3, 0.5)$) -- (xor);
\draw[arrow, blue!60!black, dashed] ($(server2.north west)+(-0.3, 0.5)$) -- (xor);

\end{tikzpicture}
\caption{Two-server batch DPF-PIR architecture. The client generates $K$ DPF key pairs (one per PBC bucket) and sends the shares to each server. Each server computes $K$ XOR sums in parallel; the client combines responses to recover all queried entries.}
\label{fig:architecture}
\end{figure}

\subsubsection{Database Generation}

The offline pipeline transforms raw Bitcoin blockchain data into two PIR-queryable databases. Each stage is a separate Rust binary in the \texttt{build} crate.

\textbf{Pipeline overview:}
\begin{enumerate}[nolistsep,leftmargin=*]
    \item \textbf{UTXO extraction:} Parse the UTXO snapshot (\texttt{dumptxoutset}) to extract ScriptPubKey hashes, full 32-byte TXIDs, vout indices, and amounts.
    \item \textbf{Chunk generation:} Group UTXOs by ScriptPubKey, serialize with VarInt encoding, and pack into fixed-size 80-byte blocks. Each block stores complete records including full TXIDs.
    \item \textbf{Index construction:} Build an index mapping each ScriptPubKey hash (20 bytes) to its chunk location and count. Each index entry is 26 bytes: 20B script hash + 4B offset + 1B chunk count + 1B flags.
    \item \textbf{Cuckoo hashing:} Organize both databases into cuckoo hash tables for PBC-compatible batch retrieval. The index uses 2-hash cuckoo with bucket size 3; the chunk database uses 3-hash cuckoo with bucket size 2. Both achieve $>$95\% load factor with zero stash.
    \item \textbf{Placement stamping:} For each index entry, determine which cuckoo hash function placed its first chunk in each of the 3 assigned groups, and encode this placement information into the 1-byte flags field (see Section~\ref{sec:placement}).
    \item \textbf{Verification:} Validate that all entries are retrievable from the cuckoo tables.
\end{enumerate}

\textbf{Dual role of cuckoo hashing.} Cuckoo hashing serves two purposes in our system. First, it provides PIR-by-keyword: given a ScriptPubKey, the client computes candidate bucket indices to locate data without scanning. Second, it implements PBC for batch PIR: when querying $m$ entries, each query maps to 3 candidate buckets; the client places queries into $K$ total buckets via cuckoo insertion, and all $K$ DPF queries execute in parallel. The index level uses 2 cuckoo hash functions with bucket size 3 (each DPF query returns 3 slots $\times$ 26~bytes = 78~bytes), while the chunk level uses 3 cuckoo hash functions with bucket size 2 (each DPF query returns 2 slots $\times$ 84~bytes = 168~bytes).

\subsubsection{Query Protocol}

A complete UTXO lookup requires two rounds of batch PIR:

\textbf{Level 1: Index lookup.} The client computes RIPEMD-160 hashes of all queried ScriptPubKeys, derives 3 candidate buckets per query via cuckoo hash functions (\texttt{splitmix64}-based), and places them into $K{=}75$ PBC buckets. For each bucket, the client generates 2 DPF key pairs (one per cuckoo hash function in the 2-hash index cuckoo) and sends one share of each to each server. The servers expand all DPF keys and return XOR sums. The client combines responses to recover 26-byte index entries containing chunk offsets, chunk counts, and placement flags.

\textbf{Level 2: Chunk retrieval.} Using the chunk locations from Level~1, the client repeats the process with $K_\text{chunk}{=}80$ PBC buckets over the chunk database. The chunk cuckoo uses 3 hash functions with bucket size 2, so the client generates 3 DPF key pairs per bucket. However, for single-chunk addresses (93.9\% of all entries), the placement optimization (Section~\ref{sec:placement}) reduces this to 1 DPF key pair per bucket. The recovered chunks contain full UTXO records: 32-byte TXIDs, vout indices, and amounts in satoshis---everything needed to construct a spending transaction.

\subsubsection{Placement Optimization}
\label{sec:placement}

For the chunk-level cuckoo table (3 hash functions, bucket size 2), each chunk can reside in one of $3 \times 3 = 9$ positions per group (3 hash functions $\times$ 2 slots is narrowed to which hash function placed it, so 3 possibilities per group). Across its 3 assigned groups, this gives $3^3 = 27$ possible placements, encodable in 5 bits.

During database construction, the build pipeline determines the actual placement of each entry's \emph{first chunk} and encodes it in the flags byte of the corresponding index entry:
$$\texttt{flags} = (1 \ll 5) \mid (h_0 + 3 \cdot h_1 + 9 \cdot h_2)$$
where $h_i \in \{0, 1, 2\}$ is the cuckoo hash function that placed the first chunk in group $i$. Bit 5 serves as a validity indicator.

At query time, the client decodes the placement bits from the recovered index entry. When issuing Level~2 queries for the first chunk, the client knows exactly which cuckoo hash function to target, reducing the DPF keys per bucket from 3 to 1. Since 93.9\% of addresses have exactly 1 chunk, this optimization applies to the vast majority of queries, yielding approximately a 67\% reduction in chunk-level communication and a 50\% reduction in server computation for those queries.

\subsubsection{Client Implementations}

\textbf{Rust CLI (\texttt{lookup\_pir}).} Command-line tool that takes ScriptPubKeys as hex, performs the two-level batch PIR protocol, and outputs decoded UTXO data.

\textbf{Web client (TypeScript).} Browser-compatible library using WebSocket connections. Implements DPF key generation, cuckoo hashing, and PBC placement in pure TypeScript. Supports Bitcoin address input (P2PKH, P2SH, P2WPKH, P2WSH, P2TR) with automatic conversion to ScriptPubKey. The web interface provides a self-contained demo at \texttt{batch\_web\_client/}.
